# Analysis of Household power consumption
Author: Ewerton Ciuffo  
Exploratory Data Analysis, class Summer 2014  
Coursera/John Hopkins

## Synopsis
This assignment uses data from the UC Irvine Machine Learning Repository, a popular repository for machine learning datasets. In particular, we will be using the "Individual household electric power consumption Data Set" made available on the course web site:

.Dataset: Electric power consumption [20Mb]


.Description: Measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.


The following descriptions of the 9 variables in the dataset are taken from the UCI web site:
1.Date: Date in format dd/mm/yyyy 
2.Time: time in format hh:mm:ss 
3.Global_active_power: household global minute-averaged active power (in kilowatt) 
4.Global_reactive_power: household global minute-averaged reactive power (in kilowatt) 
5.Voltage: minute-averaged voltage (in volt) 
6.Global_intensity: household global minute-averaged current intensity (in ampere) 
7.Sub_metering_1: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered). 
8.Sub_metering_2: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light. 
9.Sub_metering_3: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.



```{r loaddata,echo=TRUE}

setwd("C:/Users/eciuffo/github/jhopkins/eda/1")

pc <- read.table(unz("exdata_data_household_power_consumption.zip", "household_power_consumption.txt"), header=T, sep=";",stringsAsFactors = FALSE)
pc$Time <- strptime(paste(pc$Date,pc$Time), "%d/%m/%Y %H:%M:%S")

pc[, 1] <- as.Date(pc[, 1],format = "%d/%m/%Y ") 
pc.sub<-subset(pc, subset=(pc$Date>=as.Date("2007/02/01","%Y/%m/%d") & pc$Date<=as.Date("2007/02/02","%Y/%m/%d") ))

pc.sub[, 3] <- as.numeric(pc.sub[, 3])
pc.sub[, 4] <- as.numeric(pc.sub[, 4])
pc.sub[, 5] <- as.numeric(pc.sub[, 5])
pc.sub[, 6] <- as.numeric(pc.sub[, 6])
pc.sub[, 7] <- as.numeric(pc.sub[, 7])
pc.sub[, 8] <- as.numeric(pc.sub[, 8])
pc.sub[, 9] <- as.numeric(pc.sub[, 9])
```

## Plot 1- Global Active Power histogram

```{r plot1,echo=TRUE}

png(file="plot1.png",width=480,height=480,units="px")
hist(pc.sub$Global_active_power, main = "Global Active Power", xlab = "Global active Power(KiloWatts)", ylab = "Frequency", col = "red") 
dev.off()

```

## Plot 2- Global Active Power Time series

```{r plot2,echo=TRUE}

png(file="plot2.png",width=480,height=480,units="px")
plot (pc.sub$Time, pc.sub$Global_active_power,type="l",col = "blue", lwd = 1,xlab="",ylab="Global Active Power(Kilowatts)")
dev.off()

```

## Plot 3- Submetering 

```{r plot3,echo=TRUE}

png(file="plot3.png",width=480,height=480,units="px")
plot (pc.sub$Time, pc.sub$Sub_metering_1,type="l",col = "blue", lwd = 1,ylab="Energy Sub Metering",xlab="")
lines (pc.sub$Time, pc.sub$Sub_metering_2,col="red")
lines (pc.sub$Time, pc.sub$Sub_metering_3,col="green")

legend("topright",legend=c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),pch=1,col=c("blue","red","green"),lty=1)
dev.off()

```

## Plot 4 - 4 in 1  

```{r plot4,echo=TRUE}

png(file="plot4.png",width=480,height=480,units="px")
par(mfrow=c(2,2))
plot (pc.sub$Time, pc.sub$Global_active_power,type="l",col = "blue", lwd = 1,xlab="",ylab="Global Active Power(Kilowatts)")
plot (pc.sub$Time, pc.sub$Voltage,type="l",col = "blue", lwd = 1,xlab="datetime",ylab="Voltage)")
plot (pc.sub$Time, pc.sub$Sub_metering_1,type="l",col = "blue", lwd = 1,ylab="Energy Sub Metering",xlab="")
lines (pc.sub$Time, pc.sub$Sub_metering_2,col="red")
lines (pc.sub$Time, pc.sub$Sub_metering_3,col="green")
legend("topright",pch=1,col=c("blue","red","green"),legend=c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),lty=1)
plot (pc.sub$Time, pc.sub$Global_reactive_power,type="l",col = "blue", lwd = 1,xlab="datetime",ylab="Global Reactive Power(Kilowatts)")
dev.off()
```